Solving Collective Spin Hamiltonians with Decoherence

### 2018-09-04 soc_squeezing/cumulant_expansions.tex
improved text in main body and added all decoherence calculations

\newcommand{\C}{\mathcal{C}}
\newcommand{\D}{\mathcal{D}}
\newcommand{\F}{\mathcal{F}}
\newcommand{\I}{\mathcal{I}}
\renewcommand{\O}{\mathcal{O}}
\renewcommand{\S}{\mathcal{S}}

\newcommand{\N}{\mathbb{N}}
\newcommand{\Z}{\mathbb{Z}}

\newcommand{\bmu}{{\bar\mu}}
\newcommand{\floor}[1]{\left\lfloor{#1}\right\rfloor}

@@

Previous work has shown how to realize various spin-squeezing
Hamiltonians with ultracold atomic systems.  While it is
straightforward to simulate the dynamics generated by these
Hamiltonians in the context of a closed quantum system, here we try to
develop an understanding of their the squeezing behavior in the
presence of decoherence.  In particular, we will be considering
dynamics induced by the collective one-axis twisting (OAT), two-axis
twisting (TAT), and transverse field (TVF) Hamiltonians of the form
\begin{align}
  H_{\t{OAT}} = \chi S_z^2,
  &&
  H_{\t{TAT}}^{z,y} = \chi \p{S_z^2 - S_y^2},
  &&
  H_{\t{TVF}} = \chi S_z^2 + \Omega S_x,
\end{align}
for spins which decay in an uncorrelated manner at a rate $\gamma$.
The evolution of any Heisenburg operator $\O$ with a Hamiltonian $H$
is given by
\begin{align}
  \f{d}{dt} \O
  = i\sp{H,\O}_- + \gamma\D_-\p{\O},
  &&
  \D_-\p{\O} \equiv \sum_j\p{\sigma_j^+\O\sigma_j^-
    - \f12\sp{\sigma_j^+\sigma_j^-,\O}_+},
  \label{eq:EOM}
\end{align}
where $\sp{A,B}_\pm\equiv AB\pm BA$ and $j$ indexes an individual
spin.  Our measure of squeezing is
\begin{align}
  \xi^2 \equiv \f{N}{\abs{\bk{\v S}}^2}
  \min_{\uv n\perp\bk{\v S}} \bk{\p{\v S\c\uv n}^2},
  \label{eq:squeezing}
\end{align}
where $N$ is the total number of spins, $\v S$ is the collective spin
vector, and the minimization is performed over all unit vectors
$\uv n$ in the plane orthogonal to the mean spin vector $\bk{\v S}$.
This squeezing parameter is entirely determined by components
$\bk{S_\alpha}$ of the mean spin vector in addition to collective
spin-spin correlators of the form $\bk{S_\alpha S_\beta}$; our task is
therefore to compute these expectation values in all scenarios of
interest.

@@

The case of the one-axis twisting is simplest to consider, as we can
write the Hamiltonian in the form
\begin{align}
  H_{\t{OAT}} = \f14 \chi \sum_{j,k} \sigma_j^z \sigma_k^z
  = \f12 \chi \sum_{j<k} \sigma_j^z \sigma_k^z + \f14 N \chi,
\end{align}
which admits an exact solution previously worked out by Foss-Feig et
al.~in ref.~\cite{foss-feig2013nonequilibrium}.  Adapting exact
expectation values with respect to the initial state $\ket{X}$
satisfying $S_x\ket{X}=S\ket{X}$, in terms of $\mu,\nu\in\set{+1,-1}$
we find that
\begin{align}
  \bk{S_+}
  &= \sum_j\bk{\sigma_j^+}
  = \f{N}{2} e^{-\gamma t/2} \Phi\p{\chi,t}^{N-1}, \label{eq:S+_OAT} \\
  \bk{S_\mu S_z}
  &= \f12\sum_j\bk{\sigma_j^\mu\sigma_j^z}
  + \f12\sum_{j\ne k} \bk{\sigma_j^\mu \sigma_k^z} \\
  &= -\f{\mu}{2}\bk{S_\mu} + \f14 N \p{N-1} e^{-\gamma t/2}
  \Psi\p{\mu\chi,t} \Phi\p{\chi,t}^{N-2}, \\
  \bk{S_\mu S_\nu}
  &= \sum_j \bk{\sigma_j^\mu \sigma_j^\nu}
  + \sum_{j\ne k} \bk{\sigma_j^\mu \sigma_k^\nu} \\
  &= \delta_{\mu,-\nu} \p{\f{N}{2} + \mu \bk{S_z}}
  + \f14 N \p{N-1} e^{-\gamma t}
  \Phi\p{\sp{\mu+\nu}\chi,t}^{N-2}, \label{eq:SS+-_OAT}
\end{align}
where
\begin{align}
  \Phi\p{\xi,t} \equiv e^{-\gamma t/2}
  \sp{\cos\p{s_\xi t}+\f12\gamma t~\t{sinc}\p{s_\xi t}},
  &&
  \Psi\p{\xi,t} \equiv e^{-\gamma t/2}
  \p{is_\xi-\gamma/2}t~ \t{sinc}\p{s_\xi t},
\end{align}
and $s_\xi \equiv \xi + i2\gamma$.  The last ingredients we need
to compute spin squeezing at any time are $\bk{S_z}$ and $\bk{S_z^2}$.
As these operators commute with the OAT Hamiltonian, their evolution
is governed entirely by decoherence, such that
\begin{align}
  \f{d}{dt}\p{S_z^n}
  = \gamma \D_-\p{S_z^n}
  = \gamma \p{S + S_z} \sum_{k=0}^{n-1}
  \p{-1}^{n-k} { n \choose k } S_z^k,
\end{align}
where $S\equiv N/2$ throughout this work and $\D_-\p{S_z^n}$ is worked
out in Appendix \ref{sec:decoherence}.  In particular,
\begin{align}
  \f{d}{dt} S_z = -\gamma \p{S + S_z},
  &&
  \f{d}{dt}\p{S_z^2} = \gamma \p{S + S_z} \p{1 - 2S_z}.
\end{align}
The initial conditions $\bk{S_z}_{t=0}=0$ and $\bk{S_z^2}_{t=0}=S/2$
then imply
\begin{align}
  \bk{S_z} = S\p{e^{-\gamma t}-1},
  &&
  \bk{S_z^2} = S \p{1 - \f12 e^{-\gamma t}} e^{-\gamma t} + \bk{S_z}^2.
  \label{eq:Sz_OAT}
\end{align}
The expectation values in \eqref{eq:S+_OAT}-\eqref{eq:SS+-_OAT} and
\eqref{eq:Sz_OAT} are sufficient to compute the squeezing parameter
$\xi^2$ in \eqref{eq:squeezing} at any time throughout evolution of
the initial state $\ket{X}:S_x\ket{X}=S\ket{X}$ under $H_{\t{OAT}}$.

\section{Cumulant expansion and systems of equations for Heisenberg
  operators}

Unlike the one-axis twisting Hamiltonian, the two-axis twisting and
transverse-field Hamiltonians do not admit exact analytical
solutions. We will therefore solve these Hamiltonians approximately
using systems of equations for Heisenberg operators.  The basic idea
is to find equations of motion for operators of interest, and at some
point truncate high-order terms with negligible contribution to the
spin correlators of interest.  To benchmark this method, we test it
against exact one-axis twisting results in the presence of
decoherence, as well as transverse-field and two-axis twisting in the
absence of decoherence.

All collective spin Hamiltonians can be written in the form
\begin{align}
  H = \sum_{p,q,r} h_{pqr} S_+^p S_z^q S_-^r,
  \label{eq:H_general}
\end{align}
where Hermiticity requires that $h_{pqr}=h_{rqp}$.  The Hamiltonians
for one-axis twisting with a transverse field, for example, is
\begin{align}
  H_{\t{TVF}}
  = \chi S_z^2 + \Omega S_x
  = \chi S_z^2 + \f{\Omega}{2} S_+ + \f{\Omega}{2} S_-,
\end{align}
while the two-axis twisting Hamiltonians about $\p{z,y}$ and $\p{y,x}$
are
\begin{align}
  H_{\t{TAT}}^{z,y}
  = \chi \p{S_z^2 - S_y^2}
  = \chi \p{S_z^2 + \f14 S_+^2 + \f14 S_-^2 - \f12 S_+ S_- + \f12 S_z},
\end{align}
and
\begin{align}
  H_{\t{TAT}}^{y,x}
  = \chi\p{S_y^2 - S_x^2}
  = -\f{\chi}{2} \p{S_+^2 + S_-^2}.
\end{align}
The evolution of a general operator $S_+^\ell S_-^m S_z^n$ under a
Hamiltonian of the form in \eqref{eq:H_general} is given by
\begin{align}
  \f{d}{dt} \p{S_+^\ell S_z^m S_-^n}
  = i\sum_{p,q,r} h_{pqr} \sp{S_+^p S_z^q S_-^r, S_+^\ell S_z^m S_-^n}_-
  + \gamma \D_-\p{S_+^\ell S_z^m S_-^n}
  \equiv \sum_{p,q,r} T^{\ell m n}_{pqr} S_+^p S_z^q S_-^r,
  \label{eq:time_derivative}
\end{align}
where commutator and decoherence operator necessary for computing the
matrix elements $T^{\ell m n}_{pqr}$ of the time derivative operator
$d/dt$ are respectively worked out in Appendix \ref{sec:comm_general}
and \ref{sec:decoherence}.

The time derivative operator $d/dt$ will generally couple spin
operators $S_+^\ell S_z^m S_-^n$ to spin operators $S_+^p S_z^q S_-^r$
with higher weight, i.e.~with $p+q+r>\ell+m+n$.  In practice, keeping
track of this operator growth will eventually require more
computational resources than are available, so we must truncate it at
some point by taking
\begin{align}
  \f{d}{dt} \p{S_+^\ell S_z^m S_-^n}
  = \sum_{p,q,r} T^{\ell m n}_{pqr} S_+^p S_z^q S_-^r
  \to \sum_{w\p{p,q,r}<W} T^{\ell m n}_{pqr} S_+^p S_z^q S_-^r
  \label{eq:time_derivative_truncation}
\end{align}
for some cutoff $W$ and weight function $w$, e.g.~$w\p{p,q,r}=p+q+r$,
although the most optimal weight function will generally be
system-dependent.  The truncation in
\eqref{eq:time_derivative_truncation} will close the system of
differential equations defined by \eqref{eq:time_derivative}, and
allow us to solve it using standard numerical methods.  The
expectation values of collective spin operators for initial conditions
of interest are provided in Appendix \ref{sec:initial_conditions}.

Alternately, we can formally solve the differential equation in
\eqref{eq:time_derivative} and write
\begin{align}
  S_+^\ell S_z^m S_-^n
  = \left.\exp\p{t \f{d}{dt}}\right|_{t=0} \p{S_+^\ell S_z^m S_-^n}
  = \sum_{k\ge0} \f{t^k}{k!}
  \left.\f{d^k}{dt^k}\right|_{t=0} \p{S_+^\ell S_z^m S_-^n},
\end{align}
which implies
\begin{align}
  \bk{S_+^\ell S_z^m S_-^n}
  = \sum_{k\ge0} \f{t^k}{k!}
  \sum_{p,q,r} T^{\ell m n}_{pqr;k} \bk{S_+^p S_z^q S_-^r}_{t=0},
  \label{eq:time_series}
\end{align}
where the matrix elements of the $k$-th time derivative are given by
\begin{align}
  T^{\ell m n}_{pqr;0}=\delta_{\ell p}\delta_{mq}\delta_{nr},
  &&
  T^{\ell m n}_{pqr;1}\equiv T^{\ell m n}_{pqr},
\end{align}
\begin{align}
  T^{\ell mn}_{pqr;k>1}
  \equiv \sum_{\v\lambda,\v\mu,\v\nu}
  T^{\ell mn}_{\lambda_1 \mu_1 \nu_1}
  T^{\lambda_1 \mu_1 \nu_1}_{\lambda_2 \mu_2 \nu_2}
  T^{\lambda_2 \mu_2 \nu_2}_{\lambda_3 \mu_3 \nu_3}
  \cdots T^{\lambda_{k-1} \mu_{k-1} \nu_{k-1}}_{pqr}.
\end{align}
Instead of explicitly choosing a weight function that determines which
operators to keep track of, we can now truncate the series in
\eqref{eq:time_series} by taking
\begin{align}
  \bk{S_+^\ell S_z^m S_-^n}
  = \sum_{k\ge0} \f{t^k}{k!}
  \sum_{p,q,r} T^{\ell m n}_{pqr;k} \bk{S_+^p S_z^q S_-^r}_{t=0}
  \to \sum_{k<M} \f{t^k}{k!}
  \sum_{p,q,r} T^{\ell m n}_{pqr;k} \bk{S_+^p S_z^q S_-^r}_{t=0}.
\end{align}
Examining the nonzero entries of $T^{\ell mn}_{pqr;k}$ for
$k=1,2,\cdots,M-1$ now {\it tells} us which operators we need to keep
track of in order to compute the expectation value
$\bk{S_+^\ell S_z^m S_-^n}$.  Furthermore, after computing the
expectation values of all relevant operators at $t=0$, time evolution
of $\bk{S_+^\ell S_z^m S_-^n}$ takes only $\O\p{M}$ CPU time.

As a final comment, we note that the relation
$\p{S_+^\ell S_z^m S_-^n}^\dag=S_+^n S_z^m S_-^\ell$ can be exploited
to cut the number of operators we need to keep track of roughly in
two.

@@

This text makes ubiquitous use of various spin operator identities; we
collect and derive particularly useful identities here for reference.
The elementary commutation relations between single-weight spin
operators are, with $\bmu\equiv-\mu\in\set{+1,-1}$ for brevity,
\begin{align}
  \f12 \sp{\sigma_j^z,\sigma_k^\mu}_-
  &= \delta_{jk} \mu \sigma_j^\mu,
  &
  \sp{S_z,\sigma_j^\mu}_-
  &= \f12 \sp{\sigma_j^z,S_\mu}_- = \mu \sigma_j^\mu,
  &
  \sp{S_z,S_\mu}_-
  &= \mu S_\mu,
  \label{eq:comm_z_base} \\
  \sp{\sigma_j^\mu,\sigma_k^\bmu}_-
  &= \delta_{jk} \mu \sigma_j^z,
  &
  \sp{S_\mu,\sigma_j^\bmu}_-
  &= \sp{\sigma_j^\mu,S_\bmu}_- = \mu \sigma_j^z,
  &
  \sp{S_\mu,S_\bmu}_-
  &= 2 \mu S_z.
  \label{eq:comm_mu_base}
\end{align}
These relations can be used to inductively compute identities
involving powers of collective spin operators.  Working through the
possibilities, we find
\begin{align}
  S_z^m \sigma_j^\mu
  = S_z^{m-1} \sigma_j^\mu \p{\mu + S_z}
  = S_z^{m-2} \sigma_j^\mu \p{\mu + S_z}^2
  = \cdots
  = \sigma_j^\mu \p{\mu + S_z}^m,
  \label{eq:push_z_mu_Ss}
\end{align}
and
\begin{align}
  \sigma_j^z S_\mu^m
  = \p{2\mu\sigma_j^\mu + S_\mu \sigma_j^z} S_\mu^{m-1}
  = 2\mu\sigma_j^\mu S_\mu^{m-1} + S_\mu \sigma_j^z S_\mu^{m-1}
  = \cdots
  = 2\mu m\sigma_j^\mu S_\mu^{m-1} + S_\mu^m \sigma_j^z.
  \label{eq:push_z_mu_sS}
\end{align}
Summing over the single-spin index $j$ in both of these cases gives us
the purely collective-spin versions of these identities:
\begin{align}
  S_z^m S_\mu = S_\mu \p{\mu + S_z}^m,
  &&
  S_z S_\mu^m = S_\mu^m \p{\mu m + S_z},
  \label{eq:push_z_mu_single}
\end{align}
where we can repeat the process of pushing through individual
operators $\ell$ times to get
\begin{align}
  S_z^\ell S_\mu^m
  = S_z^{\ell-1} S_\mu^m \p{\mu m + S_z}
  = S_z^{\ell-2} S_\mu^m \p{\mu m + S_z}^2
  = \cdots
  = S_\mu^m \p{\mu m + S_z}^\ell.
  \label{eq:push_z_mu}
\end{align}
Similarly finding commutation relations between powers of transverse
spin operators, i.e.~$S_\mu$, turns out to be considerably more
difficult than the cases we have worked out thus far.  We therefore
save this work for Appendix \ref{sec:comm_transverse}.

\section{Commutator between powers of two transverse spin operators}
\label{sec:comm_transverse}

@@

  S_\mu^m \sigma_j^\bmu
  &= S_\mu^{m-1}\sigma_j^\bmu S_\mu + \mu S_\mu^{m-1} \sigma_j^z \\
  &= S_\mu^{m-2} \sigma_j^\bmu S_\mu^2 + \mu S_\mu^{m-2} \sigma_j^z S_\mu
  + \mu S_\mu^{m-1} \sigma_j^z \\
  &= \sigma_j^\bmu S_\mu^m
  + \mu \sum_{k=0}^{m-1} S_\mu^k \sigma_j^z S_\mu^{m-k-1}

@@

$\sp{S_\mu^m,\sigma_j^\bmu}_-$, we would like to enforce an ordering
on products of spin operators, which will allow us to only keep track
of operators that are linearly independent.  We choose to impose an
ordering with all $\sigma_j^\bmu$ operators on the left, and all
$\sigma_j^z$ operators on the right, as such an ordering will prove
convenient for the calculations in this section.  This choice of
ordering compels us to expand

@@

  S_\mu^m \sigma_j^\bmu

>>>>

\newcommand{\C}{\mathcal{C}}
\newcommand{\D}{\mathcal{D}}
\newcommand{\F}{\mathcal{F}}
\newcommand{\I}{\mathcal{I}}
\newcommand{\J}{\mathcal{J}}
\renewcommand{\O}{\mathcal{O}}
\renewcommand{\S}{\mathcal{S}}

\newcommand{\N}{\mathbb{N}}

\newcommand{\z}{\text{z}}
\newcommand{\x}{\text{x}}
\newcommand{\y}{\text{y}}
\newcommand{\Z}{\text{Z}}
\newcommand{\X}{\text{X}}
\newcommand{\bmu}{{\bar\mu}}
\newcommand{\floor}[1]{\left\lfloor{#1}\right\rfloor}

@@

\tableofcontents

\newpage

\section{Introduction}

Past work has shown how to realize spin-squeezing Hamiltonians with a
variety of ultracold atomic and trapped ion systems.  While simulating
the dynamics and squeezing behavior induced by these Hamiltonians is
straightforward in the context of a closed quantum system, here we
develop a method to simulate dynamics in the presence of decoherence.
In particular, we will consider dynamics induced by the one-axis
twisting (OAT), two-axis twisting (TAT), and transverse field (TVF)
Hamiltonians of the form
\begin{align}
  H_{\t{OAT}} = \chi S_\z^2,
  &&
  H_{\t{TAT}}^{\z,\y} = \chi \p{S_\z^2 - S_\y^2},
  &&
  H_{\t{TVF}} = \chi S_\z^2 + \Omega S_\x.
\end{align}
In order to determine the squeezing behavior of these Hamiltonians, we
will study the dynamics of Heisenburg operators $\O$, whose evolution
under a Hamiltonian $H$ is given by
\begin{align}
  \f{d}{dt} \O = i\sp{H,\O}_- + \sum_\J \gamma_\J \D\p{\J;\O},
  &&
  \D\p{\J;\O}
  \equiv \sum_{J\in\J}\p{J^\dag \O J - \f12\sp{J^\dag J,\O}_+},
  \label{eq:EOM}
\end{align}
where $\J$ is a set of jump operators with a corresponding decoherence
rate $\gamma_\J$ and $\sp{A,B}_\pm\equiv AB\pm BA$.  Our measure of
spin squeezing is
\begin{align}
  \xi^2 \equiv \f{N}{\abs{\bk{\v S}}^2}
  \min_{\uv n\perp\bk{\v S}} \bk{\p{\v S\c\uv n}^2},
  \label{eq:squeezing}
\end{align}
where $N$ is the total number of spins, $\v S$ is the collective spin
vector, and the minimization is performed over all unit vectors
$\uv n$ in the plane orthogonal to the mean spin vector $\bk{\v S}$.
This squeezing parameter is entirely determined by components
$\bk{S_\alpha}$ of the mean spin vector in addition to collective
spin-spin correlators of the form $\bk{S_\alpha S_\beta}$; our task is
therefore to compute these expectation values in all scenarios of
interest.

@@

The case of the one-axis twisting is simplest to consider, as we can
write the Hamiltonian in the form
\begin{align}
  H_{\t{OAT}} = \f14 \chi \sum_{j,k} \sigma_\z^{(j)} \sigma_\z^{(k)}
  = \f12 \chi \sum_{j<k} \sigma_\z^{(j)} \sigma_\z^{(k)} + \f14 N \chi,
\end{align}
which admits an exact solution previously worked out by Foss-Feig et
al.~in ref.~\cite{foss-feig2013nonequilibrium}.  Adapting exact
expectation values with respect to the initial state $\ket{\X}$
satisfying $S_\x\ket{\X}=S\ket{\X}$, in terms of
$\mu,\nu\in\set{+1,-1}$ we find that
\begin{align}
  \bk{S_+}
  &= \sum_j\bk{\sigma_+^{(j)}}
  = \f{N}{2} e^{-\gamma t/2} \Phi\p{\chi,t}^{N-1}, \label{eq:S+_OAT} \\
  \bk{S_\mu S_\z}
  &= \f12\sum_j\bk{\sigma_\mu^{(j)}\sigma_\z^{(j)}}
  + \f12\sum_{j\ne k} \bk{\sigma_\mu^{(j)} \sigma_\z^{(k)}} \\
  &= -\f{\mu}{2}\bk{S_\mu} + \f14 N \p{N-1} e^{-\gamma t/2}
  \Psi\p{\mu\chi,t} \Phi\p{\chi,t}^{N-2}, \\
  \bk{S_\mu S_\nu}
  &= \sum_j \bk{\sigma_\mu^{(j)} \sigma_\nu^{(j)}}
  + \sum_{j\ne k} \bk{\sigma_\mu^{(j)} \sigma_\nu^{(k)}} \\
  &= \delta_{\mu,-\nu} \p{\f{N}{2} + \mu \bk{S_\z}}
  + \f14 N \p{N-1} e^{-\gamma t}
  \Phi\p{\sp{\mu+\nu}\chi,t}^{N-2}, \label{eq:SS+-_OAT}
\end{align}
where
\begin{align}
  \Phi\p{\zeta,t} \equiv e^{-\gamma t/2}
  \sp{\cos\p{s_\zeta t}+\f12\gamma t~\t{sinc}\p{s_\zeta t}},
  &&
  \Psi\p{\zeta,t} \equiv e^{-\gamma t/2}
  \p{is_\zeta-\gamma/2}t~ \t{sinc}\p{s_\zeta t},
\end{align}
and $s_\zeta \equiv \zeta + i2\gamma$.  The last ingredients we need
to compute spin squeezing at any time are $\bk{S_\z}$ and
$\bk{S_\z^2}$.  As these operators commute with the OAT Hamiltonian,
their evolution is governed entirely by decoherence through
uncorrelated decay of individual spins, which means (see Appendix
\ref{sec:decay_single})
\begin{align}
  \f{d}{dt} S_\z = -\gamma \p{S + S_\z},
  &&
  \f{d}{dt}\p{S_\z^2} = \gamma \p{S + S_\z} \p{1 - 2S_\z},
\end{align}
where $S\equiv N/2$.  The initial conditions $\bk{S_\z}_{t=0}=0$ and
$\bk{S_\z^2}_{t=0}=S/2$ then imply
\begin{align}
  \bk{S_\z} = S\p{e^{-\gamma t}-1},
  &&
  \bk{S_\z^2} = S \p{1 - \f12 e^{-\gamma t}} e^{-\gamma t} + \bk{S_\z}^2.
  \label{eq:Sz_OAT}
\end{align}
The expectation values in \eqref{eq:S+_OAT}-\eqref{eq:SS+-_OAT} and
\eqref{eq:Sz_OAT} are sufficient to compute the squeezing parameter
$\xi^2$ defined in \eqref{eq:squeezing} at any time throughout
evolution of the initial state $\ket{\X}:S_\x\ket{\X}=S\ket{\X}$ under
$H_{\t{OAT}}$.

\section{Truncated Heisenberg operator dynamics}

Unlike the one-axis twisting Hamiltonian, the two-axis twisting and
transverse-field Hamiltonians do not admit exact analytical
solutions. We will therefore solve these Hamiltonians approximately
using alternate methods.  The basic strategy will be to find equations
of motion for operators of interest, and at some point truncate
high-order terms with negligible contribution to spin correlators of
interest.  To benchmark this method, we test it against exact one-axis
twisting results in the presence of decoherence, collective
Hamiltonian dynamics in the absence of decoherence, and exact
simulations of few-spin systems.

All collective spin Hamiltonians can be written in the form
\begin{align}
  H = \sum_{p,q,r} h_{pqr} S_+^p S_\z^q S_-^r,
  \label{eq:H_general}
\end{align}
where Hermiticity requires that $h_{pqr}=h_{rqp}$.  The Hamiltonian
for one-axis twisting with a transverse field, for example, is
\begin{align}
  H_{\t{TVF}}
  = \chi S_\z^2 + \Omega S_\x
  = \chi S_\z^2 + \f{\Omega}{2} S_+ + \f{\Omega}{2} S_-,
\end{align}
while the two-axis twisting Hamiltonians about $\p{\z,\y}$ and
$\p{\y,\x}$ are
\begin{align}
  H_{\t{TAT}}^{\z,\y}
  \propto S_\z^2 - S_\y^2
  = S_\z^2 + \f14 S_+^2 + \f14 S_-^2
  - \f12 S_+ S_- + \f12 S_\z,
\end{align}
and
\begin{align}
  H_{\t{TAT}}^{\y,\x}
  \propto S_\y^2 - S_\x^2
  = -\f12 S_+^2 - \f12 S_-^2.
\end{align}
The evolution of a general operator $S_+^\ell S_-^m S_\z^n$ under a
Hamiltonian of the form in \eqref{eq:H_general} is given by
\begin{align}
  \f{d}{dt} \p{S_+^\ell S_\z^m S_-^n}
  = i \sum_{p,q,r} h_{pqr}
  \sp{S_+^p S_\z^q S_-^r, S_+^\ell S_\z^m S_-^n}_-
  + \sum_\J \gamma_\J \D\p{\J;\O}
  \equiv \sum_{p,q,r} T^{\ell m n}_{pqr} S_+^p S_\z^q S_-^r,
  \label{eq:time_derivative}
\end{align}
where commutator and decoherence terms necessary for computing the
matrix elements $T^{\ell m n}_{pqr}$ of the time derivative operator
$d/dt$ are worked out Appendices \ref{sec:comm_general},
\ref{sec:decoherence_single}, and \ref{sec:decoherence_collective}.

The time derivative operator $d/dt$ will generally couple spin
operators $S_+^\ell S_\z^m S_-^n$ to spin operators
$S_+^p S_\z^q S_-^r$ with higher weight, i.e.~with $p+q+r>\ell+m+n$.
In practice, keeping track of this operator growth will eventually
require more computational resources than are available, so we must
truncate it at some point by taking
\begin{align}
  \f{d}{dt} \p{S_+^\ell S_\z^m S_-^n}
  = \sum_{p,q,r} T^{\ell m n}_{pqr} S_+^p S_\z^q S_-^r
  \to \sum_{w\p{p,q,r}<W} T^{\ell m n}_{pqr} S_+^p S_\z^q S_-^r
  \label{eq:time_derivative_truncation}
\end{align}
for some cutoff $W$ and weight function $w$, e.g.~$w\p{p,q,r}=p+q+r$.
The truncation in \eqref{eq:time_derivative_truncation} will close the
system of differential equations defined by
\eqref{eq:time_derivative}, and allow us to solve it using standard
numerical methods.  The initial conditions for this system of
differential equations, i.e.~expectation values of operators with
respect to states that are simple to prepare, are provided in Appendix
\ref{sec:initial_conditions}.

The truncation strategy in \eqref{eq:time_derivative_truncation} has a
few limitations:
\begin{enumerate*}[label=(\roman*)]
\item simulating a system of differential equations for a large number
  of operators can be very time-consuming,
\item the weight function $w$ must be chosen carefully, as the optimal
  weight function is generally system-dependent,
\item simulation results can only be trusted up to the time at which
  the initial value of operators $S_+^p S_\z^r S_-^q$ with weight
  $w\p{p,q,r}\ge W$ have a non-negligible contribution to operators of
  interest.
  \label{pt:limitation}
\end{enumerate*}
While the first two of these limitations might in principle be avoided
by alternate means of tracking operators, the last limitation
necessarily applies in some form to any method tracking only a subset
of {\it all} operators.  Embracing the unavoidable, we therefore try
to devise an alternate truncation strategy which is built around
\ref{pt:limitation}.

We can expand operators in a Taylor series about $t=0$ to write
\begin{align}
  S_+^\ell S_\z^m S_-^n
  = \sum_{k\ge0} \f{t^k}{k!}
  \left.\f{d^k}{dt^k}\right|_{t=0} \p{S_+^\ell S_\z^m S_-^n},
\end{align}
which implies
\begin{align}
  \bk{S_+^\ell S_\z^m S_-^n}
  = \sum_{k\ge0} \f{t^k}{k!}
  \sum_{p,q,r} T^{\ell m n}_{pqr;k} \bk{S_+^p S_\z^q S_-^r}_{t=0},
  \label{eq:time_series}
\end{align}
where the matrix elements of the $k$-th time derivative are given by
\begin{align}
  T^{\ell m n}_{pqr;0}=\delta_{\ell p}\delta_{mq}\delta_{nr},
  &&
  T^{\ell m n}_{pqr;1}\equiv T^{\ell m n}_{pqr},
\end{align}
\begin{align}
  T^{\ell mn}_{pqr;k>1}
  \equiv \sum_{\v\lambda,\v\mu,\v\nu}
  T^{\ell mn}_{\lambda_1 \mu_1 \nu_1}
  T^{\lambda_1 \mu_1 \nu_1}_{\lambda_2 \mu_2 \nu_2}
  T^{\lambda_2 \mu_2 \nu_2}_{\lambda_3 \mu_3 \nu_3}
  \cdots T^{\lambda_{k-1} \mu_{k-1} \nu_{k-1}}_{pqr}.
\end{align}
Instead of explicitly choosing a weight function that determines which
operators to keep track of, we can now truncate the series in
\eqref{eq:time_series} by taking
\begin{align}
  \bk{S_+^\ell S_\z^m S_-^n}
  = \sum_{k\ge0} \f{t^k}{k!}
  \sum_{p,q,r} T^{\ell m n}_{pqr;k} \bk{S_+^p S_\z^q S_-^r}_{t=0}
  \to \sum_{k<M} \f{t^k}{k!}
  \sum_{p,q,r} T^{\ell m n}_{pqr;k} \bk{S_+^p S_\z^q S_-^r}_{t=0}.
\end{align}
Examining the nonzero entries of $T^{\ell mn}_{pqr;k}$ for
$k=1,2,\cdots,M-1$ now {\it tells} us which operators we need to keep
track of in order to compute the expectation value
$\bk{S_+^\ell S_\z^m S_-^n}$.  Furthermore, after computing the
expectation values of all relevant operators at $t=0$, time evolution
of $\bk{S_+^\ell S_\z^m S_-^n}$ takes only $\O\p{M}$ CPU time.
Finally, the relation
$\p{S_+^\ell S_\z^m S_-^n}^\dag=S_+^n S_\z^m S_-^\ell$ can be
exploited to cut both the number of expectation values
$\bk{S_+^p S_\z^r S_-^q}_{t=0}$ and the number of matrix elements
$T^{\ell mn}_{pqr}$ which we need to compute roughly in half.

@@

The appendices in this text make ubiquitous use of various spin
operator identities; we collect and derive some basic identities here
for reference.  The elementary commutation relations between spin
operators are, with $\bmu\equiv-\mu\in\set{+1,-1}$ for brevity,
\begin{align}
  \f12 \sp{\sigma_\z^{(j)},\sigma_\mu^{(k)}}_-
  &= \delta_{jk} \mu \sigma_\mu^{(j)},
  &
  \sp{S_\z,\sigma_\mu^{(j)}}_-
  &= \f12 \sp{\sigma_\z^{(j)},S_\mu}_- = \mu \sigma_\mu^{(j)},
  &
  \sp{S_\z,S_\mu}_-
  &= \mu S_\mu,
  \label{eq:comm_z_base} \\
  \sp{\sigma_\mu^{(j)},\sigma_\bmu^{(k)}}_-
  &= \delta_{jk} \mu \sigma_\z^{(j)},
  &
  \sp{S_\mu,\sigma_\bmu^{(j)}}_-
  &= \sp{\sigma_\mu^{(j)},S_\bmu}_- = \mu \sigma_\z^{(j)},
  &
  \sp{S_\mu,S_\bmu}_-
  &= 2 \mu S_\z.
  \label{eq:comm_mu_base}
\end{align}
These relations can be used to inductively compute identities
involving powers of collective spin operators.  By pushing through one
spin operator at a time, we can find
\begin{align}
  S_\z^m \sigma_\mu^{(j)}
  = S_\z^{m-1} \sigma_\mu^{(j)} \p{\mu + S_\z}
  = S_\z^{m-2} \sigma_\mu^{(j)} \p{\mu + S_\z}^2
  = \cdots
  = \sigma_\mu^{(j)} \p{\mu + S_\z}^m,
  \label{eq:push_z_mu_Ss}
\end{align}
and
\begin{align}
  \sigma_\z^{(j)} S_\mu^m
  = \p{2\mu\sigma_\mu^{(j)} + S_\mu \sigma_\z^{(j)}} S_\mu^{m-1}
  = 2\mu\sigma_\mu^{(j)} S_\mu^{m-1} + S_\mu \sigma_\z^{(j)} S_\mu^{m-1}
  = \cdots
  = 2\mu m\sigma_\mu^{(j)} S_\mu^{m-1} + S_\mu^m \sigma_\z^{(j)}.
  \label{eq:push_z_mu_sS}
\end{align}
Summing over the single-spin index $j$ in both of these cases gives us
the purely collective-spin versions of these identities:
\begin{align}
  S_\z^m S_\mu = S_\mu \p{\mu + S_\z}^m,
  &&
  S_\z S_\mu^m = S_\mu^m \p{\mu m + S_\z},
  \label{eq:push_z_mu_single}
\end{align}
where we can repeat the process of pushing through individual $S_\z$
operators $\ell$ times to get
\begin{align}
  S_\z^\ell S_\mu^m
  = S_\z^{\ell-1} S_\mu^m \p{\mu m + S_\z}
  = S_\z^{\ell-2} S_\mu^m \p{\mu m + S_\z}^2
  = \cdots
  = S_\mu^m \p{\mu m + S_\z}^\ell.
  \label{eq:push_z_mu}
\end{align}
Similarly finding commutation relations between powers of transverse
spin operators, i.e.~$S_\mu$, turns out to be considerably more
difficult than the cases we have worked out thus far.  We therefore
save this work for Appendix \ref{sec:comm_transverse}.

\section{Commutation relations between powers of transverse spin
  operators}
\label{sec:comm_transverse}

@@

  S_\mu^m \sigma_\bmu^{(j)}
  &= S_\mu^{m-1}\sigma_\bmu^{(j)} S_\mu
  + \mu S_\mu^{m-1} \sigma_\z^{(j)} \\
  &= S_\mu^{m-2} \sigma_\bmu^{(j)} S_\mu^2
  + \mu S_\mu^{m-2} \sigma_\z^{(j)} S_\mu
  + \mu S_\mu^{m-1} \sigma_\z^{(j)} \\
  &= \sigma_\bmu^{(j)} S_\mu^m
  + \mu \sum_{k=0}^{m-1} S_\mu^k \sigma_\z^{(j)} S_\mu^{m-k-1}

@@

$\sp{S_\mu^m,\sigma_\bmu^{(j)}}_-$, we would like to enforce an
ordering on products of spin operators, which will allow us to only
keep track of operators that are linearly independent.  We choose to
impose an ordering with all $\sigma_\bmu^{(j)}$ operators on the left,
and all $\sigma_\z^{(j)}$ operators on the right, as such an ordering
will prove convenient for the calculations in this section.  This
choice of ordering compels us to expand

@@

  S_\mu^m \sigma_\bmu^{(j)}
